A Theory for Migration of Information Systems

### 2022-09-10 2022 Migration/articleMigration.tex
wat probeerseltjes...

\subsection{Information Systems}
\label{sct:Information Systems}
	Information systems process information from many different users,
	all with their own tasks and responsibilities,
	and all working from different locations and on different moments.
	This collective use serves a purpose which we loosely call "the business".
	As a consequence, the data in an information system changes continually.
	To preserve meaning, users are trying to maintain semantic constraints on the data.
	So, let us first define the constituent notions role and constraint before defining information systems proper.

@@

	A \define{constraint} is a restriction on a dataset, to be enforced by a role.
	The purpose of constraints is to ensure the semantics of a dataset, i.e. the preservation of meaning as data changes.
	Each constraint $u\in\rules$ has a signature $\signat{u}$, which is a pair of concepts.
	Each constraint $u$ also has a function $\id{viol}_u:\powerset{\atoms\times\rels\times\atoms}\ \rightarrow\ \powerset{\atoms\times\atoms}$.
	We call the set $\viol{\triples}{u}$ the \define{violation set} of constraint $u$ in population $\triples$.
	We call every element $\pair{a}{b}\in\viol{\triples}{u}$ a \define{violation} of constraint $u$ in $\triples$.
	We say that a constraint \define{is satisfied} if its violation set is empty (i.e. if $\viol{\triples}{u}$ is the empty set).
	The set of constraints is called $\rules$.

@@

	We can now define the notion of information system.
	Let $\roles$ be a set of roles,
	let $\rules$ be a set of constraints, 
	let $\maintain$ be a surjective relation $\roles\times\rules$,
	and let $\dataset$ be a dataset.
\begin{definition}[information system]
	\label{def:information system}
	\item An information system is a tuple $\la\roles,\rules,\id{maintain},\dataset\ra$\\
	with dataset $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$
	in which $o\ \maintain\ u$ means that role $o$ must keep the set $\viol{\triples}{u}$ empty.
\end{definition}
	If a constraint $u$ has violations, i.e. $\viol{\triples}{u}$ is not empty, there is a role whose job it is to remove those violations.
	The relation $\maintain$ being surjective means that all users together are keeping the violation set empty.
	Which means that all constraints are being kept satisfied in the collective effort of users, even if the data changes.
	In another section we will give some options on how users can keep rules satisfied.

>>>>

\subsection{Information Systems}
\label{sct:Information Systems}
	Information systems process information from many different users,
	all with their own tasks and responsibilities,
	and all working from different locations and on different moments.
	This collective use serves a purpose which we loosely call ``the business''.
	As a consequence, the data in an information system changes continually.
	To preserve meaning, users are trying to maintain semantic constraints on the data.
	So, let us first define the constituent notions role and constraint before defining information systems proper.

@@

	A \define{constraint} is a restriction on a dataset, to be enforced by a role.
	The purpose of constraints is to ensure the semantics of a dataset, i.e. the preservation of meaning as data changes.
	Each constraint $u\in\rules$ has a signature $\signat{u}$, which is a pair of concepts.
	Each constraint $u$ also has a function $\id{viol}_u:\powerset{\atoms\times\rels\times\atoms}\ \rightarrow\ \powerset{\atoms\times\atoms}$.
	We call the set $\viol{\triples}{u}$ the \define{violation set} of constraint $u$ in population $\triples$.
	We call every element $\pair{a}{b}\in\viol{\triples}{u}$ a \define{violation} of constraint $u$ in $\triples$.
	We say that a constraint \define{is satisfied} if its violation set is empty.
	The set of constraints is called $\rules$.

@@

	We can now define the notion of information system.
	Let $\roles$ be a set of roles,
	let $\rules$ be a set of constraints, 
	let $\maintain$ be a surjective relation $\roles\times\rules$,
	and let $\dataset$ be a dataset.
\begin{definition}[information system]
	\label{def:information system}
	\item An information system $\infsys$ is a tuple $\la\roles,\rules,\maintain,\dataset\ra$\\
	with dataset $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$
	in which $o\ \maintain\ u$ means that role $o$ must keep the set $\viol{\triples}{u}$ empty.
\end{definition}
	If a constraint $u$ has violations, i.e. $\viol{\triples}{u}$ is not empty, there is a role whose job it is to remove those violations.
	The relation $\maintain$ being surjective means that all users together are keeping the violation set empty.
	Which means that all constraints are being kept satisfied in the collective effort of users, even if the data changes.
	In another section we will give some options on how users can keep rules satisfied.

\subsection{Migration}
	To migrate information system $\infsys$ to $\infsys'$, we take the disjoint union $\infsys\sqcup\infsys'$ of both information systems
	and add rules to describe the migration.
	$\infsys$ is the existing system, which contains data to be preserved.
	$\infsys'$ is the new system, which contains no or hardly any data.
	It does, however, contain the concepts, relations, rules and roles of the new system.

	We use two datasets: $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ and $\dataset'=\la\atoms',\concepts',\inst',\isa',\rels',\triples'\ra$.
	Before doing so, let us first define the disjoint union of two information systems.
\begin{definition}[disjoint union of datasets]
	\begin{eqnarray}
		\dataset\sqcup\dataset'&=&\la\atoms\sqcup\atoms',\ \concepts\sqcup\concepts',\ \inst\sqcup\inst',\ \isa\sqcup\isa',\ \rels\sqcup\rels',\ \triples\sqcup\triples'\ra\\
		\atoms\sqcup\atoms'&=&\{(a,0)|\ a\in\atoms\}\ \cup\ \{(a',1)|\ a'\in\atoms'\}\\
		\concepts\sqcup\concepts'&=&\{(c,0)|\ c\in\concepts\}\ \cup\ \{(c',1)|\ c'\in\concepts'\}\\
		\inst\sqcup\inst'&=&\begin{array}[t]{@{}l}
							  \{((a,0),(A,0))|\ (a,A)\in\inst\}\ \cup\\
			                  \{((a',1),(A',1))|\ (a',A')\in\inst'\}
							\end{array}\\
		\isa\sqcup\isa'&=&\begin{array}[t]{@{}l}
							  \{((A,0),(B,0))|\ A\isa B\}\ \cup\\
			                  \{((A',1),(B',1))|\ A'\isa' B'\}
							\end{array}\\
		\rels\sqcup\rels'&=&\{(r,0)|\ r\in\rels\}\ \cup\ \{(r',1)|\ r'\in\rels'\}\\
		\triples\sqcup\triples'&=&\begin{array}[t]{@{}l}
			                        \{\triple{(a,0)}{(r,0)}{(b,0)}|\ \triple{a}{r}{b}\in\triples\}\ \cup\\
			                        \{\triple{(a',1)}{(r',1)}{(b',1)}|\ \triple{a'}{r'}{b'}\in\triples'\}
								   \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[$\doteq$]
	\begin{eqnarray}
		\declare{nm}{A}{B}\doteq\declare{nm'}{A'}{B'}&\Leftrightarrow&\id{nm}=\id{nm'}\wedge A=A'\wedge B=B'
	\end{eqnarray}
\end{definition}
\begin{definition}[migration dataset]
	\begin{eqnarray}
		&&\{ {\tt ENFORCE\ }r'{\tt := }r| r\in\rels, r'\in\rels', r\doteq r'\}\\
		&\cup&\begin{array}[t]{@{}l}
				\{ {\tt ENFORCE\ }\declare{nm'}{A'}{B'}{\tt := }\declare{nm}{A}{C};\ident{B}\\
				|\ \begin{array}[t]{@{}l}
					\declare{nm}{A}{C}\in\rels\wedge B\isa C\ \wedge\\
					\declare{nm'}{A'}{B'}\in\rels'\wedge B'\not\isa C\ \wedge\\
					\declare{nm}{A}{B}\doteq\declare{nm'}{A'}{B'}\}
				   \end{array}
			  \end{array}\\
		&\cup&\begin{array}[t]{@{}l}
				\{ {\tt ENFORCE\ }\declare{nm'}{A'}{C'}{\tt := }\ident{A};\declare{nm}{B}{C}\\
				|\ \begin{array}[t]{@{}l}
					\declare{nm}{B}{C}\in\rels\wedge A\isa B\ \wedge\\
					\declare{nm'}{A'}{C'}\in\rels'\wedge A'\not\isa B\ \wedge\\
					\declare{nm}{A}{C}\doteq\declare{nm'}{A'}{C'}\}
				   \end{array}
			  \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[disjoint union of information systems]
	\begin{eqnarray}
		\infsys\sqcup\infsys'&=&\la\roles\sqcup\roles',\rules\sqcup\rules',\maintain\sqcup\maintain',\dataset\sqcup\dataset'\ra\\
		\roles\sqcup\roles'&=&\{(r,0)|\ r\in\roles\}\ \cup\ \{(r',1)|\ r'\in\roles'\}\\
		\rules\sqcup\rules'&=&\{(r,0)|\ r\in\rules\}\ \cup\ \{(r',1)|\ r'\in\rules'\}\\
		\maintain\sqcup\maintain'&=&\begin{array}[t]{@{}l}
							  \{((o,0),(u,0))|\ (o,u)\in\maintain\}\ \cup\\
			                  \{((o',1),(u',1))|\ (o',u')\in\maintain'\}
							\end{array}
	\end{eqnarray}
\end{definition}
